Variation in use of East Asian Late Paleolithic weapons: A study of tip cross-sectional area of stemmed points from Korea

### 2022-11-21 analysis/paper/paper.qmd
Edited discussion

@@

```{r}
#| label: fig-tukey-eco-region
#| fig-cap: "."
#| fig-width:   5
#| fig-height:  4

@@

ggplot(eco_tur_df, aes(row.names(eco_tur_df))) + 
  geom_boxplot(aes(middle=diff, upper=upr, lower=lwr, ymin=lwr, ymax=upr), stat="identity")

@@

@@

We then located individual assemblages on maps that show eco-regional boundaries to explore spatial patterns.

We used two maps from the previous studies that represents different zones based on vegetation and geography [@yi2010vegetation; @yim1975distribution; @lee2008relationship]. Our results show that stemmed points are only located in certain zones, Central Temperature Zone (CT) or Central inland (CI) (@fig-gis-tcsa-sites), for example. We observed that TCSA values are stonely associated with the zones (Vegetation zones: `apastats::describe.Anova(eco_lgm_test, 1, 2)`, Ecoregion zones: `apastats::describe.Anova(for_lgm_test, 1, 2)`). However, there is no clear functional pattern throughout the zones (@fig-gis-tcsa-sites panel B and D). There is a wide variation in TCSA regardless of the stemmed points density in each zone.

Tukey

@@

Our results indicate that javelin tips and stabbing spear tips are the most probable ballistic uses for stemmed points. There are a few stemmed points in our sample that may have been used as dart tips and un-poisoned arrow tips, but none of them have TCSA values in the range of poison arrow tips (@fig-tcsa-all-sp). But in general, we found a wide range of TCSA in our sample, indicating that stemmed points played a diverse role in forager toolkits. TCSA can be impacted by other factors such as raw materials and portability. @eren2022plains explain that smaller TCSA could be the result of pursuing production economy and transport efficiency. We explore the relationship between TCSA and raw material types and weight in order to examine these factors. We observed no clear pattern for raw materials (@fig-tcsa-raw-materials) but weight (@fig-artifact-size-cluster). Interestingly, weight is correlated with raw materials (@fig-tcsa-size). Therefore we could assume that raw materials also associated with TCSA value. We found that different clusters of weight are matched with different types of weapon tips (@fig-artifact-size-cluster). The TCSA varies from assemblages to assemblages without any discernible temporal pattern in functional use (@fig-tcsa-radio-carbon). But the stemmed points mainly used before LGM (@fig-tcsa-radio-carbon-box). We found that stemmed points are only located in certain ecoregional zones such as Central Temperature zone without any clear pattern in function (@fig-gis-tcsa-sites). Both temporal and spatial analyses show that stemmed points were used for a particular environment, not so much harsh as LGM.

@@

THIS DISCUSSION IS INCOMPLETE, WE NEED TO EXPLORE THESE RELATIONSHIPS ABOUT GEOGRAPHICAL VARIATION AND TEMPORAL VARIATION FURTHER, AND IN A NEW WAY THAT IS NOT THE SAME AS THE QSR PAPER, E.G. FOCUS ON THE LGM AS I MENTIONED IN OUR LAST MEETING.

SPATIAL ANAlysis @prates2022changes provde strong relationship between the temporal density and spatial distribution of megafauna and Fishtail projectile points, and proposed that this weapon was designed and used for megafauna hunting, contributing to their extinction by exploring changes in projectile point technology, as well as the body mass of potential megafaunal prey, and show that Fishtails were strongly related to the largest extinct megafaunal species.

>>>>

# explore artifacts in certain zone, related to Tukey's test
tukey_veg <- points_in_veg_zones_all %>% 
  filter(ecoregion == "ST") %>% 
  dplyr::select(ecoregion, Assemblage, min_tcsa, raw_materials_top, age_ka)

@@

```{r}
#| label: fig-tukey-region
#| fig-cap: "."
#| fig-width:   5
#| fig-height:  4

@@

eco_tur_box <- ggplot(
  data = eco_tur_df,
  mapping = aes(
    x = row.names(eco_tur_df),
    lower = lwr,
    middle = diff,
    upper = upr,
    ymin = lwr,
    ymax = upr
    )
  ) +
  geom_boxplot(stat = "identity")+
  xlab("Vegetation zones")+
  theme(axis.text.x = element_text(angle = 45, size = 6)) +
  theme(axis.text.y = element_text(size = 6))

@@

for_tur_df <- as.data.frame(for_tur$`as.factor(Zone)`)

for_tur_box <- ggplot(
  data = for_tur_df,
  mapping = aes(
    x = row.names(for_tur_df),
    lower = lwr,
    middle = diff,
    upper = upr,
    ymin = lwr,
    ymax = upr
    )
  ) +
  geom_boxplot(stat = "identity") +
  xlab("Eco-region zones") +
  theme(axis.text.x = element_text(angle = 45, size = 6)) +
  theme(axis.text.y = element_text(size = 6))

# ggplot(for_tur_df) +
#   geom_errorbar(aes(
#     x = row.names(for_tur_df),
#     ymin = lwr,
#     ymax = upr
#     ))

plot_grid(eco_tur_box, for_tur_box, labels = c('A', 'B'))

@@

In order to explore spatial patterns as well as the relationship between the function of the tools and the environment, we located individual assemblages on maps that depict eco-regional boundaries. Our results show that stemmed points are only located in certain zones, Central Temperature Zone (CT) of vegetation map or Central inland (CI) (@fig-gis-tcsa-sites) of eco-region map. We observed that TCSA values are strongely associated with the zones (Vegetation zones: `apastats::describe.Anova(eco_lgm_test, 1, 2)`, Ecoregion zones: `apastats::describe.Anova(for_lgm_test, 1, 2)`). To explore the TCSA distribution within the eco-regional zones, we excluded zones in the ecoregion map that have no stemmed points (@fig-gis-tcsa-sites panel D). We found no clear functional pattern throughout the zones, showing a wide variation in TCSA regardless of the density of stemmed point in each zone (@fig-gis-tcsa-sites panel B and D).

We then computed Tukey's for comparison tests between the eco-regional zones. Our results show that ST and CT, and SWT-ST have significantly different while SWT and CT are close to each other (@fig-tukey-region panel A). We found no significant difference within the zones of eco-region map (@fig-tukey-region panel B).

@@

Our results indicate that javelin tips and stabbing spear tips are the most probable ballistic uses for stemmed points. There are a few stemmed points in our sample that may have been used as dart tips and un-poisoned arrow tips, but none of them have TCSA values in the range of poison arrow tips (@fig-tcsa-all-sp). But in general, we found a wide range of TCSA in our sample, indicating that stemmed points played a diverse role in foraging toolkits. TCSA can be impacted by other factors such as raw materials and portability. @eren2022plains explain that smaller TCSA could be the result of pursuing production economy and transport efficiency. We explore the relationship between TCSA and raw material types and weight in order to examine these factors. We observed no clear pattern for raw materials (@fig-tcsa-raw-materials) but weight (@fig-artifact-size-cluster), though weight is correlated with raw materials (@fig-tcsa-size). Therefore we could assume that raw materials also associated with TCSA value. We found that different clusters of weight are matched with different types of weapon tips (@fig-artifact-size-cluster). The TCSA varies from assemblages to assemblages without any discernible temporal pattern in functional use (@fig-tcsa-radio-carbon). But the stemmed points mainly used before LGM (@fig-tcsa-radio-carbon-box). We found that stemmed points are only located in certain eco-regional zones such as Central Temperature zone without any clear pattern in function (@fig-gis-tcsa-sites). Both temporal and spatial analyses show that stemmed points were used for a particular environment, not so much harsh as LGM. @prates2022changes claimed that fishtail projectile points in South America was used for megafauna hunting and contributed to their extinction by proving strong relationship between the temporal density and spatial distribution of megafauna and the projectile points. Our results support this claim by showing the dense distribution of stemmed point in certain environment.

@@
